---
title: "homework2"
author: "Trevor Caughlin"
date: "January 20, 2018"
output:
  word_document: default
  html_document: default
---
https://ms.mcmaster.ca/~bolker/emdbook/book.pdf 
Likelihood chapter beginning on page 228 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

homework 1:
Run a logistic regression on these data;
what is your interpretation of the slope?
*make it a real data set*
present the confidence interval

homework 2: 
run a binomial regression on these data

present the confidence intervals


1. use `glm` to analyze the dataset "SEEDLING_SURVIVAL.csv" These data represent the annual survival rates of 1435 tagged tree seedlings in Huai Kha Khaeng Wildlife Sanctuary, Thailand. Run separate glm models for:
a) effect of height on seedling survival
b) effect of light on seedlings survival

As output, include a sentence with parameter interpretation for both models: "The baseline survival was..., The effect of height on seedling survival was ..."
include two separate plots with `curve` to overlay model predictions on data
Also include a confindence intervals using `confint` function

From your results, which predictor variable has a stronger impact on output?

1. plot the data. Note that plotting the proportion of eggs is necessary, since the eggs and mosquito larvae both vary

2. Which is a better model to fit the data?
polynomial

just do in word
1.44+-0.19*x+-0.21*x^2+0.04*x^3

10*x*exp(-2*x)

a. use `curve` with `add=T` to overlay curves with both parameters on the data

b. use `dbinom` to calculate the likelihood of both models (with given parameter values). Note: be sure to use `log=T` for taking `sum` 
of dbinom output

or some other function? use dbinom, log and sum

homework 3:
power analysis
use rbinom and rnorm
glm for both

note this will involve a for loop

step 1: create an empty vector for for loop to fill in
Here, the object `empty vector` is just `NA` repeated 100 times. We are going to fill in these values using the for loop.

```{r forloop}
empty_vector<-rep(NA,times=100)

for(i in 1:100) {
    empty_vector[i]<-rnorm(n=1,mean=0,sd=1)
  }
```
In this example, `i` is a counter for each iteration of the for loop (from 1 to 100). At each of i iterations, R is filling in `empty_vector` with a draw from `rnorm`


how many samples do you need to
(a) recapture the slope value?

(b) ensure a p-value for slope parameter <0.05

output should include:
a) numeric values
b) plot of sample size (x-axis) and power (either p-value or bias in parameters) on y-axis


